---
title: Report
output_format: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    warning = FALSE,
    message = FALSE
)
```

```{r libraries}
library(tidyverse)
library(gt)
library(targets)
```

# Introduction

TODO

# Dataset

TODO. Check `data/README.md`.

# Model

\begin{align}
\text{mean rate of rates} &= \frac{\text{max score}}{\text{max duration}}
= \frac{100 \text{ points}}{15 \text{ minutes}} \\
\text{rate of rates} &= \frac{1}{\text{mean rate of rates}} \\
\\
\text{rate}^{\text{blue}} &\sim \text{Exponential}(\text{rate of rates}) \\
\text{rate}^{\text{red}} &\sim \text{Exponential}(\text{rate of rates}) \\
\\
\text{ttp}^{\text{blue}}_i &= \text{time}^{\text{blue}}_i - \text{time}^{\text{blue}}_{i-1} \\
\text{ttp}^{\text{red}}_i &= \text{time}^{\text{red}}_i - \text{time}^{\text{red}}_{i-1} \\
\\
\text{ttp}^{\text{blue}}_i &\sim \text{Exponential}(\text{rate}^{\text{blue}}) \\
\text{ttp}^{\text{red}}_i &\sim \text{Exponential}(\text{rate}^{\text{red}})
\end{align}

# Prior model analysis

```{r}
tar_read(prior_model_rate_plot)
```

```{r}
tar_read(prior_model_score_plot)
```

```{r}
tar_read(prior_model_ttp_vs_time_plot)
```

```{r}
tar_read(prior_model_ttp_plot)
```

# Exploratory data analysis

```{r}
tar_read(observed_score_plot)
```

```{r}
tar_read(observed_ttp_vs_time_plot)
```

```{r}
tar_read(observed_ttp_plot)
```

# Results

```{r}
tar_read(model_rate_plot)
```

```{r}
tar_read(model_contrast_plot)
```

```{r}
tar_read(model_score_plot)
```

```{r}
tar_read(model_ttp_vs_time_plot)
```

```{r}
tar_read(model_ttp_plot)
```

# Conclusion

TODO
