---
title: Bayesian halo
output: distill::distill_article
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    warning = FALSE,
    message = FALSE
)
```

```{r libraries}
library(tidyverse)
library(gt)
library(targets)
```

# Too long, didn't read (aka abstract)

TODO.

# Introduction

If you are like me, you have been playing halo infinite a lot in the past weeks. One game mode that I particularly like is big team battle slayer in which 2 teams of 16 players face each other in a battlefield. Each time one player of team A defeats a player of team B, team A earns a point. The first team to achieve 100 points first wins the game.

Now if you are really like me, you have also been wondering about how to statistically model this game mode. Specifically, I was interested in calculating the probability of my team winning at each timepoint of a match. Intuitively, I know that there is a 50%-50% chance of winning when the game starts. If the score is 20 - 10 for my team, I also expect to have a higher although not extreme probability of winning. But what about when the score is 90 - 80? From my experience this game is a certain win, but how should I quantify this probability? This is the question my work sets to answer using bayesian statistics.

# Dataset

TODO. Check `data/README.md`.

# Exploratory data analysis

```{r}
tar_read(observed_score_plot)
```

```{r}
tar_read(observed_ttp_vs_score_plot)
```

```{r}
tar_read(observed_ttp_plot)
```

# Model

\begin{align}
\text{mean rate of rates} &= \frac{\text{max score}}{2 / 3 \cdot \text{max duration}}
= \frac{100 \text{ points}}{10 \text{ minutes}} \\
\text{rate of rates} &= \frac{1}{\text{mean rate of rates}} \\
\\
\text{rate}^{\text{blue}} &\sim \text{Exponential}(\text{rate of rates}) \\
\text{rate}^{\text{red}} &\sim \text{Exponential}(\text{rate of rates}) \\
\\
\text{ttp}^{\text{blue}}_i &= \text{time}^{\text{blue}}_i - \text{time}^{\text{blue}}_{i-1} \\
\text{ttp}^{\text{red}}_i &= \text{time}^{\text{red}}_i - \text{time}^{\text{red}}_{i-1} \\
\\
\text{ttp}^{\text{blue}}_i &\sim \text{Exponential}(\text{rate}^{\text{blue}}) \\
\text{ttp}^{\text{red}}_i &\sim \text{Exponential}(\text{rate}^{\text{red}})
\end{align}

# Prior model analysis

```{r}
tar_read(prior_model_rate_plot)
```

```{r}
tar_read(prior_model_score_plot)
```

```{r}
tar_read(prior_model_ttp_vs_score_plot)
```

```{r}
tar_read(prior_model_ttp_plot)
```

# Results

```{r}
tar_read(model_rate_plot)
```

```{r}
tar_read(model_contrast_plot)
```

```{r}
tar_read(model_score_plot)
```

```{r}
tar_read(model_ttp_vs_score_plot)
```

```{r}
tar_read(model_ttp_plot)
```

# Conclusion

TODO
